---
title: "ggfast: Extensible package for data reduction in aid of visualization"
author: "Malachi Phillips"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

# Motivation

Deciphering data is hard. Deciphering *lots* of data is *very* hard. The initial motivation for writing a package such as _ggfast_ was in order to enable users to better control their data, rather than having the data control them. _ggfast_ relies on several different 'data reduction' and 'predictive' techniques that all fall into the guise of classical statistics/machine learning concepts. What's most powerful in the types of analysis that _ggfast_ relies on is the ability to reduce data: taking in large sets of multivariate data and reducing it down to just a few clusters, or just a few principle components. In doing these analyses provided in _ggfast_, one hopes to be able to separate the signal from the noise within a data set, in order to reach real, data-driven inferences from these analyses.

# Theoretical Background for Analyses

_ggfast_ relies on a few of the following different analyses:
  - Kmeans clustering
  - Principle Component Analysis (PCA)
  - Multi-linear regression (MLR)
  
Although the means through which each analysis works differ, the key points of all of these analyses are the same:

  1) Reduce the size of the data, whether into clusters, principle components, or a single equation
  2) Provide a natural error estimator for how much error occurs in the data reduction, whether through cluster gap errors, less-constributing principle components, or linear residuals
  3) Offer a predictive tool, whether through representing a data point by its nearest cluster, making an inference from the equation of the principal component, or relying on the model provided by a multi-linear regression.
  
In this regard, each of these analyses has the attractive offer shown in (1)-(3), but may achieve it in a different way.

## Kmeans clustering

Kmeans clustering aims to take our multivariate data set and reduce them into $k$ centroids that may then be used to predict values based on the inferred value of the closest centroid.
For example, suppose 19 clusters are generated from some huge, multivariate data set, and we wish to infer a measured variable $f$ from the clusters. We then, taking the information of the new point, assign it to the closest cluster, say cluster #5. From here, we infer the new data's value of $f$ based on the value of $f$ from cluster 5.
In particular, kmeans clustering is attractive as it provides a way of predicting `qualitative, categorical variables`. In this regard, kmeans clustering is the only effective tool for the prediction of a qualitative, categorical variable in _ggfast_. The other analyses are ill-adapted to analyze categorical variables.

The kmeans clustering algorithm may be described by the following pseudo code:

1) Initialize *cluster centroids* $\mu_1,mu_2,...,\mu_k$ randomly, where $k$ is the number of clusters we are considering.

2) Repeat until convergence/maximum number of iterations: {
  For every *i* indexing the observations, set to the closest cluster, ie. $c^{(i)}= arg min_j |x^{(i)}-\mu_j|^2$. This sets the observation to the nearest cluster available.
  For every *j*, re-calculate the cluster means for each $\mu_j$ based on what data is currently assigned to each cluster.

## Principal Component Analysis (PCA)

Principal component analysis (PCA) is a form of analysis that seeks to _reduce_ the dimensionality of a multivariate data set, hoping to reduce, say $p=12$ variables into two or three main components which are made up of the other variables. Furthermore, it is also possible using a PCA analysis to see what, if any, variables may be singly contributing to the variability of a data set. In this regard, a PCA analysis can be _elucidating_ into the nature of the data set, and similarly has both predictive power (in the equations drawn up from the principal components), and an error estimator (in the error taken from not considering several components).

The basis of PCA lies in the eigenvector-eigenvalue problem, namely:

$$
Ax=\lambda x
$$

Where we are seeking to find all $(\lambda_i,x_i)$ pairs. Then by sorting the eigenvector/eigenvalues by the largest magnitude eigenvalues, we can calculate the amount of variability explained in an individual component by:

$$
\dfrac {\lambda_j} {\sum_{i} \lambda_{i}}
$$
Further, the principal components have an equation of the form:

$$
Y_i = e_i'(x-\mu)
$$
Often, a PCA analysis can explain 95% of the variability with just the top two principal components. However, this isn't always the case -- many times, PCA analyses offer little to no insight. 

## Multiple Linear Regression

Linear regression models are formulated to estimate the linear dependence of variables. For example, a p-dimensional data set would be approximated as follows:

$$
y_i = \beta_0 + \beta_1 x_{1,i} + ... \beta_p x_{p,i}
$$

It also follows that:

$$
Y = X\beta + \epsilon
$$
This can also be summarized as a series of vector-matrix operations, with $X$ denoting the design matix, which is the data matrix (minus the response) with a column of 1's appended to the leftmost column:

$$
\beta = (X'X)^{-1}X'Y
$$
Of course, $\beta$ is merely a point-estimate for the population parameter $\beta$, and so there is some need to do bootstrapping to obtain distributional information about $\beta$. Therefore, bootstrapping is implemented in the package.

One particular problem that arises when bootstrapping is the issue of finding an ill-conditioned matrix. Once this occurs, there are a few options available:

  1. try to approximately solve the system by perturbing the matrix in a small way that does not affect the solution too drastically, known as _preconditioning_
  
  2. simply ignore the cases when the matrix is singular, and re-sample
  
  3. By using 2., get an initial distribution for $\beta$, then re-run the bootstrap with the knowledge of the approximate, biased distribution of $\beta$. Whenever a singular matrix is encountered, merely use the distribution to generation a random value for $\beta$ at that point. This can be iteratively tried again, but typically, $n=3$ tries is sufficient.

In the function `bootreg`, the first choice is chosen. In particular, a technique known as the _preconditioned conjugate gradient method_ is employed to approximately solve, in an iterative fashion, the equation: $(X'X)\beta=X'Y$, which is equivalent to finding $\beta=(X'X)^{-1}X'Y$, which involves an expensive inverse calculation. A full discussion of the _preconditioned conjugate gradient method_ is beyond the scope of this vignette, but in short, the psuedocode below sums it up:

```{}
# For solving Ax=b, M is a 'preconditioner matrix'
ro<-b-Ax
zo<-inv(m) ro
po<-zo
k<-0
while(k < maximum_allowable_iterations){
  alpha_k <- rk'zk/(pk'Apk)
  xk+1<-xk+alpha_k pk
  rk+1<-rk-alpha_k A pk
  if rk+1 is sufficiently small, exit loop
  z_k+1<-inv(M) rk+1
  beta_k <- zk+1'rk+1/zk'rk
  pk+1<-zk+1 + beta_k p_k
  k<-k+1
}
return xk+1 as result
```

where $M$ is a _magical preconditioner matrix_, which is generally generated by an _incomplete cholesky factorization_, which itself may be generated using the following pseudo code:

```{}
function a = ichol(a)
	n = size(a,1);

	for k=1:n
		a(k,k) = sqrt(a(k,k));
		for i=(k+1):n
		    if (a(i,k)!=0)
		        a(i,k) = a(i,k)/a(k,k);            
		    endif
		endfor
		for j=(k+1):n
		    for i=j:n
		        if (a(i,j)!=0)
		            a(i,j) = a(i,j)-a(i,k)*a(j,k);  
		        endif
		    endfor
		endfor
	endfor

    for i=1:n
        for j=i+1:n
            a(i,j) = 0;
        endfor
    endfor            
endfunction
```

The power of multilinear regression, like the other methods, is that it offers:

  - an error estimator, in the form of $\epsilon_i = |y_i-\hat{y_i}|$
  - predictive power, in the form of $\hat{Y}=X\beta$

With this information, we are able to move onto the implementation details.

One other important point is in the creation of a weighted least squares formulation, which is comparable to multilinear regression, but rather the point estimate for $\beta$ is:

$$
\beta = (X'WX)^{-1}X'WY
$$

where $W$ represents a diagonal matrix that is the inverse of the expected variances within each term. This becomes particularly important whenever there exists an error that is not normally distributed with a constant variance, as is the assumption in ordinary linear regression. Typically, weighted least squares regression models are more well formed, but only if the weights are chosen in a logical, consistent way.

# Functions in ggfast

## Kmeans clustering

The (current) functionality of _ggfast_ included is:

(1) Cluster reduction/write to file function.

Writes to a .csv file the output of a cluster reduction, up to a given k.

Implementation:

```{r, eval=FALSE}
#'
#' Reduce the amount of data through using kmeans clustering algorithm
#' Call the in R kmean cluster function in order to reduce a pool of data into a few, clustered points
#'
#' @param x Data in the form of a data frame or data matrix, for safety, please use data.matrix(...) as a wrapper
#' @param k Number of clusters to form
#' @param fileName Name of file you wish to save a .csv formatted file to
#' @param iter.max Number of iterations to use at a maximum (default = 10)
#' @param nstart How many random sets to be chosen
#' @param algorithm What algorithm to use (default="Lloyd").
#'                  choices include: "Hartigan-Wong", "LLoyd", "MacQueen", and "Forgy"
#' @param trace logical or integer number, currently only used in the default method ("Hartigan-Wong"):
#'  if positive (or true), tracing information on the progress of the algorithm is produced.
#'  Higher values may produce more tracing information.
#'
#'  @examples
#'  reduce_cluster(data.matrix(iris), 10, "myFile.csv")
#'  reduce_cluster(df, 15, "myManyIterationsFile.csv", iter.max=1000, algorithm="MacQueen")
#'
reduce_cluster<-function(x, k, fileName, iter.max = 10, nstart = 1, algorithm="Lloyd", trace = FALSE)
{
  # perform cluster analysis to find means of each cluster group
  cl<-kmeans(x,k,iter.max=iter.max,nstart=nstart,algorithm=algorithm,trace=trace)
  reduced_data<-cl$centers

  # write output data in the form of a .csv file
  write.csv(reduced_data, fileName)

}
```

Examples:

```{r, eval=FALSE}
reduce_cluster(data.matrix(iris), 10, "output.csv")
```

(2) The ability to visualize the reduced data, on top of the original data:

Implementation:

```{r, eval=FALSE}
#'
#' Visualize the amount of data reduction in the first component from using clustering
#' This verifies that the clustering analysis can be of some use for data reduction
#'
#' @param x Data in the form of a data frame or data matrix, for safety, please use data.matrix(...) as a wrapper
#' @param k Number of clusters to form
#'
#' @examples
#' visualize_reduction(data.matrix(iris), 19)
visualize_reduction<-function(x,k)
{
  cl<-kmeans(x,k)
  reduced_data<-cl$centers
  windows()
  plot(x)
  points(reduced_data,col="Blue", add=TRUE)
}
```

Examples:

```{r, eval=FALSE}
visualize_reduction(data.matrix(iris),15)
```

(3) A detailed analysis on the impact of the choice of k, along with several other useful analyses.

Implementation:

```{r, eval=FALSE}
#'
#' Perform the types of anaylsis which do a reduction on the data set,
#' finding the optimum value of k
#'
#' NOTE: by default, this will perform the analysis on the scaled data
#'
#' @param x Data in the form of a data frame or data matrix, for safety, please use data.matrix(...) as a wrapper
#' @param kmax (=10) Maximum number of clusters to form
#' @param elbowPlotFileName (="elbowPlot.jpeg"), name of file for elbow plot output
#' @param silhouetteFileName (="silhouette.jpeg"), name of file for silhouette plot output
#' @param gapFileName (="gap.jpeg"), name of file for output of gap statistics plot
#' @param iter.max (=100) Maximum number of iterations
#' @param nstart How many random sets to be chosen
#' @param algorithm What algorithm to use (default="Lloyd").
#'                  choices include: "Hartigan-Wong", "LLoyd", "MacQueen", and "Forgy"
#' @param trace logical or integer number, currently only used in the default method ("Hartigan-Wong"):
#'  if positive (or true), tracing information on the progress of the algorithm is produced.
#'  Higher values may produce more tracing information.
#' @param debugOutput (=FALSE) Boolean value for whether or not to output silhouette plots for
#' each and every cluster value used. This generates a TON of output! YOU HAVE BEEN WARNED!
#' @param debugOutFileNameBase (="sil") Base file name for debug output, if enabled.
#'
#' @examples
#' cluster_k_analysis(data.matrix(iris), kmax=19)
#' cluster_k_analysis(data.matrix(iris), kmax=19, iter.max=200, debugOutput=T)
cluster_k_analysis<-function(x,
                             kmax=10,
                             elbowPlotFileName = "elbowPlot.jpeg",
                             silhouetteFileName = "silhouette.jpeg",
                             gapFileName = "gap.jpeg",
                             iter.max = 100,
                             nstart = 1,
                             algorithm="Lloyd",
                             trace = FALSE,
                             debugOutput=FALSE,
                             debugOutFileNameBase="sil")
{
  # normalize the data
  x<-scale(x)
  wss<-function(k){
    kmeans(x,k,iter.max,nstart,algorithm,trace)$tot.withinss
  }

  # Compute from k = 1 to k = n
  n = dim(x)[1]-1
  k.values <- 2 : kmax

  # extract wss for 2-n clusters
  wss_values<-purrr::map_dbl(k.values,wss)

  # Output the 'elbow plot' approach to determining the ideal
  # number of clusters to use in this type of analysis
  jpeg(elbowPlotFileName)

  # make a plot
  plot(k.values, wss_values,
       type="b", pch = 19, frame = FALSE,
       xlab="Number of clusters K",
       ylab="Total within-clusters sum of squares")

  dev.off()

  # function to compute average silhouette for k clusters
  avg_sil <- function(k) {
    km.res <- kmeans(x, centers = k, nstart = 25)
    ss <- cluster::silhouette(km.res$cluster, dist(x))
    if(debugOutput){
      name=paste(debugOutFileNameBase, k, ".jpeg", sep="")
      jpeg(name)
      plot(ss)
      dev.off()
    }
    mean(ss[,3])
  }

  avg_sil_values <- purrr::map(k.values, avg_sil)

  jpeg(silhouetteFileName)
  plot(k.values, avg_sil_values,
       type = "b", pch = 19, frame = FALSE,
       xlab = "Number of clusters K",
       ylab = "Average Silhouettes")
  dev.off()

  gap_stat<-cluster::clusGap(x,FUN=kmeans, nstart=nstart, K.max=kmax, B=50)
  jpeg(gapFileName)
  plot(gap_stat, xlab="Number of clusters k")
  dev.off()

  list(GapStatistics = gap_stat)
}
```

Examples:

```{r, eval=FALSE}
cluster_k_analysis(data.matrix(iris), kmax=19, iter.max=200, debugOutput=T)
```

## Principal Component Analysis

The current functioanlity of _ggfast_ currently includes:

(1) Determining greatest factors, and % of variability explained with $k$ factors:

Implementation:

```{r, eval=FALSE}
#'
#' Determine greatest data factors
#' @param x Data set to be analyzes
#' @param k Number of components considered
#'
#' @examples
#' determine_greatest_contributors(data.matrix(iris), 2)
#'
determine_greatest_contributors<-function(x, k)
{
  x<-scale(x)
  S<-cov(x)
  eig_res<-eigen(S)

  lambda_sum = sum(eig_res$value)
  n = length(eig_res$value)
  prop<-matrix(NA, nr=n)
  for(i in 1 : n){
    prop[i]=eig_res$value[i]/lambda_sum
  }

  err<-1-sum(prop[1:k])

  list(EigenValues = eig_res$value, EigenvalueSum= lambda_sum, Proportions=prop, Error=err, EigenVectors=eig_res$vectors)
}
```

Examples:

```{r, eval=FALSE}
determine_greatest_contributors(data.matrix(iris), 2)
```

(2) Run a PCA analysis, showing the components, as well as a built in clustering analysis.

Implementation:

```{r, eval=FALSE}
#'
#' After calling whatever means of cluster reductions, now do a PCA
#' on the already reduced data set.
#'
#' @param x Data set to be analyzed
#' @param k Number of components considerd
#' @param barPlotFileName (="barPlotComp.jpeg") file name used to store bar plot output
#'
PCAPlot<-function(x, k, barPlotFileName="barPlotComp.jpeg")
{
  unscaled<-x # keep a local copy for later
  x<-scale(x)
  pca<-determine_greatest_contributors(x,k)
  PC<-x%*% pca$EigenVectors
  Component<-colnames(x)
  Proportion<-pca$Proportions
  Errors<-1-pca$Proportions
  df<-data.frame(Component,Proportion,Errors)

  theme<-ggplot2::theme_set(cowplot::theme_cowplot()) + ggplot2::theme(legend.position="none")
  g<-ggpubr::ggbarplot(df, x="Component", y="Proportion", fill="Component",palette="jco",ggtheme=theme, sort.val="desc")
  p<-ggpubr::ggbarplot(df, x="Component", y="Errors", fill="Component", palette="jco",ggtheme=theme, sort.val="desc")
  graph<-cowplot::plot_grid(g,p)
  cowplot::save_plot(barPlotFileName, graph, base_aspect_ratio=3.0)

  # Now, in the case of k = 1,2, or 3, we can visualize the data and apply SLR on it
  data<-cbind(Component, Proportion)
  data<-data[order(-Proportion),]

  # Take the top k parts only
  names<-data[(1:k),1]
  #print(data)
  print(names)
  reduced_scaled_data_set<-subset(x, select=names) # Additionally have the option of outputting the data

  reduced_unscaled_data_set<-subset(x, select=names)
  df<-reshape2::melt(reduced_unscaled_data_set)

  if(k == 2){
    windows()
    plot(reduced_unscaled_data_set,
         main="Reduced Data Set Scatter Plot",
         xlab=colnames(reduced_unscaled_data_set)[1],
         ylab=colnames(reduced_unscaled_data_set)[2])

    # After performing this analysis, we can also do a cluster analysis
    # Typically, k = 5 clusters is an appropriate enough choice for now
    k = 5
    clus<-kmeans(reduced_unscaled_data_set,centers=k)
    windows()
    cluster::clusplot(reduced_unscaled_data_set, clus$cluster)
  }

  # Additional plot -- look at the clusplot for all of the data
  k = 5
  clus<-kmeans(x, centers=k)
  windows()
  cluster::clusplot(x, clus$cluster)

  # We may now extend this into three dimensions instead
  kdf<-kmeans(x,4)
  newdf<-data.frame(x, K=kdf$cluster)
  pcdf<-princomp(x,cor=T,score=T)
  summary(pcdf)
  rgl::plot3d(pcdf$scores, col=newdf$K)

}
```

Examples:

```{r, eval=FALSE}
PCAPlot(data.matrix(iris), 2)
```

## Multiple Linear Regresion

(1) We need a function that implements the ability to solve the equation $(X'X)\beta=X'Y$.

Implementation:

```{r, eval=FALSE}
#'
#' Point estimate for beta, used in linear regression
#'
#' Notice that this is done by solving the linear system:
#' (X'X) beta = X'Y
#'
#' Rather than using a direct method, ie.
#'
#' beta = (X'X)^(-1)X'Y
#'
#' We shall instead use a package known as 'pcig' that solves
#' a pre-conditioned conjugate gradient formulation, and is more
#' numerically stable whenever the matrix approaches singularity.
#'
#' @param X Design matrix
#' @param Y Response vector
#' @examples
#' betah(matrix(c(1,2,3,4,5,6), nr=3, byrow=TRUE),
#'       matrix(c(1,2,3), nr=3))
#'
betah<-function(X,Y){
  if(!require("pcg")) install.packages("pcg")
  pcg::pcg(t(X)%*%X, t(X)%*%Y)
}
```


Examples:

```{r, eval=FALSE}
betah(matrix(c(1,2,3,4,5,6), nr=3, byrow=TRUE),
      matrix(c(1,2,3), nr=3))
```


(2) A method that appends 1 column of 1's to convert a data matrix into a design matrix

Implementation:

```{r, eval=FALSE}
#'
#' Prepare the data matrix by appending a column of 1's top it
#' @param df Data frame, minus the response
#' @examples
#' append_design_matrix(matrix(c(1,2,3,4,5,6), nr=2, byrow=TRUE))
#'
#' df<-iris
#' df<-df[,2:dim(df)[2]] # remove first column, using it as the response vector
#' append_design_matrix(df)
#'
append_design_matrix<-function(df){

  # strip off non numeric data
  df<-df[,sapply(df, is.numeric)]
  X<-matrix(NA_real_, nr=dim(df)[1], nc=dim(df)[2]+1, byrow=TRUE)
  X[,1] = 1.0
  for(i in 1 : dim(df)[2]){
    for(j in 1 : dim(df)[1]){
      X[j,i+1] = df[j,i]
    }
  }
  return (X)
}
```


Examples:

```{r, eval=FALSE}
append_design_matrix(matrix(c(1,2,3,4,5,6), nr=2, byrow=TRUE))

df<-iris
df<-df[,2:dim(df)[2]] # remove first column, using it as the response vector
append_design_matrix(df)
```


(3) A method to do bootstrap to find confidence intervals on $\beta$ values.

Implementation:

```{r, eval=FALSE}
#'
#' Perform boot strapping
#' @param Y response vector
#' @param X design matrix
#' @param iter(=1) number of iterations to run bootstrap
#' @param alpha (=0.05) significance level for generating confidence intervals
#' @examples
#' df<-iris
#'
#' # scrape off all non-numeric data first
#' df<-df[,sapply(df, is.numeric)]
#' df
#' X<-df[,-1]
#' Y<-as.matrix(df[,1])
#' X
#' Y
#' dim(X)
#' dim(Y)
#' X<-append_design_matrix(X)
#' dim(X)
#' bh=bootreg(Y,X,1000) # works just fine! Whoo!
#' bh
bootreg=function(Y,X,iter=1, alpha = 0.05)
{
  n=dim(X)[1]
  # n is the number of rows
  nb=dim(X)[2]
  # nb is the number of columns and is equal to number of betas to be estimated
  beta.mat=matrix(NA_real_,nrow = nb,ncol=iter)
  index.mat=matrix(NA_real_,nrow=n,ncol=iter,byrow=FALSE)
  for(i in 1:iter)
  {
    index.mat[,i] = sample(1:n,n,replace=TRUE)
    #index.mat will be a matrix of random indices
    X.iter<-matrix(NA_real_, nrow=n, ncol = nb, byrow=TRUE)
    Y.iter<-matrix(NA_real_, nrow=n, ncol = 1, byrow = TRUE)
    for(j in 1 : n){
      X.iter[j,] = X[index.mat[j,i],]
      Y.iter[j,1] = Y[index.mat[j,i]]
    }
    beta.mat[,i]= betah(X.iter, Y.iter)
    #beta.mat is a matrix filled with bootstrap beta estimates
  }
  layout(matrix(1:nb, nr=nb,nc=1,byrow=TRUE))
  CIs = c();
  for(i in 1:nb)
  {
    j=i-1
    CI = quantile(beta.mat[i,], c(alpha, 1-alpha))
    CIs = c(CIs, CI)
    hist(beta.mat[i,],freq=FALSE,
         ylab="Density"  , main=substitute(beta[j]),
         xlab = substitute(beta[j]))
    mtext(paste("CI = (", round(CI[1],4), ",",round(CI[2],4),")" ), cex=0.6)
  }


  list(bh=beta.mat, Y=Y,X=X, Confidence_Intervals = CIs)
}

```

Examples:

```{r, eval=FALSE}
df<-iris
# scrape off all non-numeric data first
df<-df[,sapply(df, is.numeric)]
df
X<-df[,-1]
Y<-as.matrix(df[,1])
X
Y
dim(X)
dim(Y)
X<-append_design_matrix(X)
dim(X)
bh=bootreg(Y,X,1000) # works just fine! Whoo!
bh
```


(4) Predict information for a given design matrix subset based on the results of boostrapping.

Implementation:


```{r, eval=FALSE}
# Some other cools ideas: add in a predictor method

#'
#' Predict the next value using the same multilinear regression model
#'
#' @param X design matrix
#' @param model the result of running bootreg on the data set, ie. a great big collection
#' of different betas, upon which the column mean is used to get the most accurate possible
#' estimate for beta
#' @examples
#' # predict a singular value, given the description of the linear model and the design matrix of the new points
#' df<-iris
#'
#' # scrape off all non-numeric data first
#' df<-df[,sapply(df, is.numeric)]
#' df
#' X<-df[,-1]
#' Y<-as.matrix(df[,1])
#' X
#' Y
#' dim(X)
#' dim(Y)
#' X<-append_design_matrix(X)
#' dim(X)
#' bh=bootreg(Y,X,1000) # works just fine! Whoo!
#' data.matrix(rowMeans(bh$bh))
#' predictor(matrix(c(1, 2, 2, 3), nr=1), bh)
#'
predictor<-function(X, model){
  betas = model$bh

  # Use the point estimate for the mean of the betas
  beta_means<-data.matrix(rowMeans(betas))

  # Y = Xbeta
  X%*%beta_means

}
```


Examples:


```{r, eval=FALSE}
df<-iris
# scrape off all non-numeric data first
df<-df[,sapply(df, is.numeric)]
df
X<-df[,-1]
Y<-as.matrix(df[,1])
X
Y
dim(X)
dim(Y)
X<-append_design_matrix(X)
dim(X)
bh=bootreg(Y,X,1000) # works just fine! Whoo!
data.matrix(rowMeans(bh$bh))
predictor(matrix(c(1, 2, 2, 3), nr=1), bh)

```


And the equivalent methods for weighted least squares:

```{r, eval=FALSE}
#'
#' Point estimate for beta, used in linear regression
#'
#' Notice that this is done by solving the linear system:
#' (X'WX) beta = X'WY
#'
#' Rather than using a direct method, ie.
#'
#' beta = (X'WX)^(-1)X'WY
#'
#' We shall instead use a package known as 'pcig' that solves
#' a pre-conditioned conjugate gradient formulation, and is more
#' numerically stable whenever the matrix approaches singularity.
#'
#' @param X Design matrix
#' @param Y Response vector
#' @param W weightt matrix
#' @examples
#' betah(matrix(c(1,2,3,4,5,6), nr=3, byrow=TRUE),
#'       matrix(c(1,2,3), nr=3).
#'       matrix(c(1.7,0,0,0.5), nr =2, byrow=TRUE))
#'
betah_wls<-function(X,Y,W){
  if(!require("pcg")) install.packages("pcg")
  pcg::pcg(t(X)%*%W%*%X, t(X)%*%W%*%Y)
}

#'
#' Perform boot strapping
#' @param Y response vector
#' @param X design matrix
#' @param W weight matrix
#' @param iter(=1) number of iterations to run bootstrap
#' @param alpha (=0.05) significance level for generating confidence intervals
#' @examples
#' df<-iris
#'
#' # scrape off all non-numeric data first
#' df<-df[,sapply(df, is.numeric)]
#' df
#' X<-df[,-1]
#' Y<-as.matrix(df[,1])
#' X
#' Y
#' dim(X)
#' dim(Y)
#' X<-append_design_matrix(X)
#' dim(X)
#' bh=bootreg(Y,X,1000) # works just fine! Whoo!
#' bh
bootreg_wls=function(Y,X,W,iter=1, alpha = 0.05)
{
  n=dim(X)[1]
  # n is the number of rows
  nb=dim(X)[2]
  # nb is the number of columns and is equal to number of betas to be estimated
  beta.mat=matrix(NA_real_,nrow = nb,ncol=iter)
  index.mat=matrix(NA_real_,nrow=n,ncol=iter,byrow=FALSE)
  for(i in 1:iter)
  {
    index.mat[,i] = sample(1:n,n,replace=TRUE)
    #index.mat will be a matrix of random indices
    X.iter<-matrix(NA_real_, nrow=n, ncol = nb, byrow=TRUE)
    Y.iter<-matrix(NA_real_, nrow=n, ncol = 1, byrow = TRUE)
    for(j in 1 : n){
      X.iter[j,] = X[index.mat[j,i],]
      Y.iter[j,1] = Y[index.mat[j,i]]
    }
    beta.mat[,i]= betah(X.iter, Y.iter, W)
    #beta.mat is a matrix filled with bootstrap beta estimates
  }
  layout(matrix(1:nb, nr=nb,nc=1,byrow=TRUE))
  CIs = c();
  for(i in 1:nb)
  {
    j=i-1
    CI = quantile(beta.mat[i,], c(alpha, 1-alpha))
    CIs = c(CIs, CI)
    hist(beta.mat[i,],freq=FALSE,
         ylab="Density"  , main=substitute(beta[j]),
         xlab = substitute(beta[j]))
    mtext(paste("CI = (", round(CI[1],4), ",",round(CI[2],4),")" ), cex=0.6)
  }
  
  
  list(bh=beta.mat, Y=Y,X=X, Confidence_Intervals = CIs)
}

```
